### 2017-06-04 sheets/talk-6/sheet-6.tex
Sheet 6: Fix two typos

\begin{exercise}[Disjoint union]
  \begin{itemize}
    \item[(a)]For an index set $I$ and a family of sets $(X_i)_{i \in I}$, refine the arbitrary disjoint union $\coprod_{i \in I}{X_i}$.
    \item[(b)]Show that this arbitrary disjoint union is indeed the arbitrary coproduct in $\Set$.
  \end{itemize}
\end{exercise}

@@

\begin{exercise}[Coequalizer of sets]
  In this exercise we will first define \emph{reflexive symmetric transitive closures} and will define the coequalizer of $\Set$.
  \begin{itemize}
    \item[(a)]First we will show that the reflexive, symmetric and transitive closure of a relation $R$ is the \emph{least} equivalence relation containing $R$.
      Let $R$ be a relation over a set $X$ and $S$ be the reflexive, symmetric and transitive closure of $R$. Show following lemmas:
      \begin{itemize}
        \item[(i)]Show $\forall n\in\nat: R_n$ and $S$ are reflexive and symmetric.
        \item[(ii)]Show $\forall n\in\nat: R \subseteq R_n$ and $R \subseteq S$.
        \item[(iii)]Show that $R_n$ is \emph{monotone}, i.\,e. $\forall n,m\in\nat: n\le m \Rightarrow R_n \subseteq R_m$.
        \item[(iv)]Now show that $S$ is transitive.
          \begin{hint}
            Let $m\in\nat$. Show $\forall x_1, x_2, x_3: (x_1, x_2) \in S \Rightarrow (x_2, x_3) \in R_m \Rightarrow (x_1, x_3) \in S$ via $\nat$-induction.
          \end{hint}
        \item[(v)]Now show that $R$ is the \emph{least} equivalence relation containing $R$, i.\,e.
          $$\forall R': \text{$R'$ is a equivalence relation} \land R \subseteq R' \Rightarrow S \subseteq R'.$$
      \end{itemize}
    \item[(b)]Now we will define the coequalizer of $\Set$.
      Let $X, Y \in \Set, s, t \from X \to Y$.
      We make the following definitions:
      \begin{alignat*}{2}
        R    &:=& \set{(s(a), t(a)) | a\in Y} \\
        \sim &:=& \text{reflexive, symmetric and transitive closure of $R$} \\
        C    &:=& Y / \sim
      \end{alignat*}
      Note that $[y] := \set{ y' \in Y | (y, y') \in \sim}$ is the equivalence class of $y$ and $Y / \sim := \set{ [y] | y \in Y}$ is the set of all equivalence classes of $\sim$.
      Also note that no equivalence class can be empty and for $[y]$ we call $y$ a representant of this class.
      We define $i \from Y \to E$ with $i(y) := [y]$.
      \begin{itemize}
        \item[(i)]Show that the following diagramm commutes:
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
          }$
        \item[(ii)]Let
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^f & A
          }$
          be a cummutative diagramm. Show that $f$ respects equivalence, i.\,e. $\forall (y_1, y_2) \in \sim: f(y_1) = f(y_2)$.
          \begin{hint}
            ``Destruct'' $(y_1, y_2) \in \sim$ and do a $\nat$-induction.
          \end{hint}
        \item[(iii)]Now define $\bar{f} \from C \to A$, such that the following diagramm commutes and show that there can only be one:
          \[ \xymatrix{
            & & A  \\
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[ru]^f \ar[rd]_i \\
            & & C \ar@{.>}[uu]|{\bar{f}}
          } \]
        \item[(iv)]Show that the fork
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
          }$.
          is the coequalizer of
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
          }$.
      \end{itemize}
  \end{itemize}
\end{exercise}

>>>>

\begin{exercise}[Disjoint union]
  \begin{itemize}
    \item[(a)]For an index set $I$ and a family of sets $(X_i)_{i \in I}$, define the arbitrary disjoint union $\coprod_{i \in I}{X_i}$.
    \item[(b)]Show that this arbitrary disjoint union is indeed the arbitrary coproduct in $\Set$.
  \end{itemize}
\end{exercise}

@@

\begin{exercise}[Coequalizer of sets]
  In this exercise we will first define \emph{reflexive symmetric transitive closures} and will define the coequalizer of $\Set$.
  \begin{itemize}
    \item[(a)]First we will show that the reflexive, symmetric and transitive closure of a relation $R$ is the \emph{least} equivalence relation containing $R$.
      Let $R$ be a relation over a set $X$ and $S$ be the reflexive, symmetric and transitive closure of $R$. Show following lemmas:
      \begin{itemize}
        \item[(i)]Show $\forall n\in\nat: R_n$ and $S$ are reflexive and symmetric.
        \item[(ii)]Show $\forall n\in\nat: R \subseteq R_n$ and $R \subseteq S$.
        \item[(iii)]Show that $R_n$ is \emph{monotone}, i.\,e. $\forall n,m\in\nat: n\le m \Rightarrow R_n \subseteq R_m$.
        \item[(iv)]Now show that $S$ is transitive.
          \begin{hint}
            Let $m\in\nat$. Show $\forall x_1, x_2, x_3: (x_1, x_2) \in S \Rightarrow (x_2, x_3) \in R_m \Rightarrow (x_1, x_3) \in S$ via $\nat$-induction.
          \end{hint}
        \item[(v)]Now show that $R$ is the \emph{least} equivalence relation containing $R$, i.\,e.
          $$\forall R': \text{$R'$ is a equivalence relation} \land R \subseteq R' \Rightarrow S \subseteq R'.$$
      \end{itemize}
    \item[(b)]Now we will define the coequalizer of $\Set$.
      Let $X, Y \in \Set, s, t \from X \to Y$.
      We make the following definitions:
      \begin{alignat*}{2}
        R    &:=& \set{(s(a), t(a)) | a\in Y} \\
        \sim &:=& \text{reflexive, symmetric and transitive closure of $R$} \\
        C    &:=& Y / \sim
      \end{alignat*}
      Note that $[y] := \set{ y' \in Y | (y, y') \in \sim}$ is the equivalence class of $y$ and $Y / \sim := \set{ [y] | y \in Y}$ is the set of all equivalence classes of $\sim$.
      Also note that no equivalence class can be empty and for $[y]$ we call $y$ a representant of this class.
      We define $i \from Y \to E$ with $i(y) := [y]$.
      \begin{itemize}
        \item[(i)]Show that the following diagramm commutes:
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
          }$
        \item[(ii)]Let
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^f & A
          }$
          be a cummutative diagramm. Show that $f$ respects equivalence, i.\,e. $\forall (y_1, y_2) \in \sim: f(y_1) = f(y_2)$.
          \begin{hint}
            ``Destruct'' $(y_1, y_2) \in \sim$ and do a $\nat$-induction.
          \end{hint}
        \item[(iii)]Now define $\bar{f} \from C \to A$, such that the following diagramm commutes and show that there can only be one:
          \[ \xymatrix{
            & & A  \\
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[ru]^f \ar[rd]_i \\
            & & C \ar@{.>}[uu]|{\bar{f}}
          } \]
        \item[(iv)]Show that the fork
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
          }$
          is the coequalizer of
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
          }$.
      \end{itemize}
  \end{itemize}
\end{exercise}
